---
title: "📊 Dynamic Report Template — Databricks"
author: "Your Name"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    number_sections: true
    code_folding: show
---

```{css, echo=FALSE}
/* ============================================================
   CUSTOM CSS STYLING
   ============================================================ */

/* General body styling */
body {
  font-family: 'Segoe UI', Arial, sans-serif;
  font-size: 15px;
  background-color: #f5f7fa;
  color: #2c3e50;
}

/* Page header */
h1.title {
  color: #1a73e8;
  font-size: 2.2em;
  border-bottom: 3px solid #1a73e8;
  padding-bottom: 10px;
}

/* Section headings */
h1, h2, h3 {
  color: #2c3e50;
}

h2 {
  border-left: 5px solid #1a73e8;
  padding-left: 10px;
  margin-top: 30px;
}

/* Info / highlight box */
.info-box {
  background-color: #e8f4fd;
  border-left: 5px solid #1a73e8;
  padding: 12px 16px;
  margin: 15px 0;
  border-radius: 4px;
}

/* Warning box */
.warning-box {
  background-color: #fff8e1;
  border-left: 5px solid #f39c12;
  padding: 12px 16px;
  margin: 15px 0;
  border-radius: 4px;
}

/* Success box */
.success-box {
  background-color: #e8f5e9;
  border-left: 5px solid #27ae60;
  padding: 12px 16px;
  margin: 15px 0;
  border-radius: 4px;
}

/* KPI metric cards */
.kpi-row {
  display: flex;
  gap: 16px;
  margin: 20px 0;
  flex-wrap: wrap;
}

.kpi-card {
  flex: 1;
  min-width: 160px;
  background: white;
  border-radius: 8px;
  padding: 18px 22px;
  box-shadow: 0 2px 8px rgba(0,0,0,0.08);
  text-align: center;
}

.kpi-card .kpi-value {
  font-size: 2em;
  font-weight: bold;
  color: #1a73e8;
}

.kpi-card .kpi-label {
  font-size: 0.85em;
  color: #888;
  margin-top: 4px;
}

/* Code chunk styling */
pre {
  background-color: #f8f9fa;
  border: 1px solid #e0e0e0;
  border-radius: 5px;
}
```

```{r setup, include=FALSE}
# ============================================================
# SETUP CHUNK — Runs silently, loads all libraries
# ============================================================
knitr::opts_chunk$set(
  echo    = TRUE,      # Show code in output
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  fig.width  = 9,
  fig.height = 5,
  fig.align  = "center"
)

# --- Install packages (run once, then comment out) ----------
# install.packages(c("knitr","kableExtra","DT","ggplot2","plotly",
#                    "dplyr","leaflet","htmlwidgets","flexdashboard",
#                    "scales","RColorBrewer"))

# --- Load libraries -----------------------------------------
library(knitr)
library(kableExtra)
library(DT)
library(ggplot2)
library(plotly)
library(dplyr)
library(leaflet)
library(htmlwidgets)
library(scales)
library(RColorBrewer)

# --- Create sample dataset ----------------------------------
set.seed(42)
sales_data <- data.frame(
  Month     = month.abb,
  Region    = rep(c("North", "South", "East", "West"), length.out = 12),
  Sales     = round(runif(12, 50000, 200000), 0),
  Target    = round(runif(12, 80000, 180000), 0),
  Growth    = round(runif(12, -10, 30), 1),
  Customers = sample(100:500, 12)
)

sales_data <- sales_data %>%
  mutate(
    Achievement = round((Sales / Target) * 100, 1),
    Status      = ifelse(Sales >= Target, "✅ Met", "❌ Missed")
  )
```

---

# Introduction {.tabset}

<div class="info-box">
📌 **About this Report:** This template demonstrates every major component you need for
dynamic HTML reports in Databricks using R Markdown. Each section below covers a
different package with working, copy-paste-ready examples.
</div>

## KPI Summary Cards

KPI cards are built with plain HTML + CSS — no package needed.

```{r kpi-cards, echo=FALSE}
total_sales   <- sum(sales_data$Sales)
avg_growth    <- mean(sales_data$Growth)
top_region    <- sales_data %>% group_by(Region) %>%
                   summarise(s = sum(Sales)) %>% arrange(desc(s)) %>% slice(1) %>% pull(Region)
met_target    <- sum(sales_data$Status == "✅ Met")
```

<div class="kpi-row">
  <div class="kpi-card">
    <div class="kpi-value">`r paste0("$", formatC(total_sales, format="f", digits=0, big.mark=","))`</div>
    <div class="kpi-label">Total Sales</div>
  </div>
  <div class="kpi-card">
    <div class="kpi-value">`r paste0(round(avg_growth,1), "%")`</div>
    <div class="kpi-label">Avg Growth</div>
  </div>
  <div class="kpi-card">
    <div class="kpi-value">`r top_region`</div>
    <div class="kpi-label">Top Region</div>
  </div>
  <div class="kpi-card">
    <div class="kpi-value">`r met_target` / 12</div>
    <div class="kpi-label">Targets Met</div>
  </div>
</div>

## Data Preview

```{r data-preview}
# Quick look at the dataset used throughout this report
head(sales_data, 6)
```

---

# kableExtra — Styled Static Tables

`kableExtra` turns plain R tables into beautifully formatted HTML tables.

## Basic Styled Table

```{r kable-basic}
sales_data %>%
  select(Month, Region, Sales, Target, Status) %>%
  kable(
    caption = "📋 Monthly Sales vs Target",
    format  = "html",
    format.args = list(big.mark = ",")
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = FALSE,
    position   = "center",
    font_size  = 13
  ) %>%
  row_spec(0, bold = TRUE, background = "#1a73e8", color = "white") %>%
  column_spec(3, bold = TRUE, color = "white",
              background = ifelse(sales_data$Sales >= sales_data$Target,
                                  "#27ae60", "#e74c3c"))
```

## Grouped Headers + Footnote

```{r kable-grouped}
sales_data %>%
  select(Month, Sales, Target, Achievement, Growth) %>%
  kable(
    caption = "📊 Performance Summary with Grouped Headers",
    col.names = c("Month", "Sales ($)", "Target ($)", "Achievement (%)", "Growth (%)"),
    format = "html"
  ) %>%
  kable_styling(bootstrap_options = c("striped","hover"), full_width = TRUE) %>%
  add_header_above(c(" " = 1,
                     "Financial Metrics" = 2,
                     "Performance Metrics" = 2)) %>%
  row_spec(0, bold = TRUE, background = "#2c3e50", color = "white") %>%
  footnote(
    general = "Data generated for illustration purposes.",
    number  = c("Achievement = Sales / Target × 100",
                "Growth compared to previous period")
  )
```

---

# DT — Interactive Searchable Tables

`DT` creates fully interactive tables — sortable, searchable, paginated, exportable.

## Full-Featured DT Table

```{r dt-full}
datatable(
  sales_data,
  caption  = "🔍 Interactive Sales Table — Search, Sort, Export",
  rownames = FALSE,
  filter   = "top",           # Column-level filters
  extensions = "Buttons",
  options = list(
    pageLength = 6,
    scrollX    = TRUE,
    dom        = "Bfrtip",   # B=Buttons, f=search, r=processing, t=table, i=info, p=pagination
    buttons    = list("copy", "csv", "excel", "pdf", "print"),
    columnDefs = list(
      list(className = "dt-center", targets = "_all")
    )
  )
) %>%
  formatCurrency(columns = c("Sales", "Target"), currency = "$", digits = 0) %>%
  formatStyle(
    "Achievement",
    background = styleColorBar(range(sales_data$Achievement), "#1a73e8"),
    backgroundSize    = "100% 80%",
    backgroundRepeat  = "no-repeat",
    backgroundPosition= "center"
  ) %>%
  formatStyle(
    "Status",
    color = styleEqual(c("✅ Met", "❌ Missed"), c("#27ae60", "#e74c3c")),
    fontWeight = "bold"
  )
```

<div class="info-box">
💡 **Tip:** Use the **Export buttons** above to download this table as CSV, Excel, or PDF.
The column filter boxes let you filter each column independently.
</div>

---

# ggplot2 — Static Charts

`ggplot2` is the most powerful static charting library in R.

## Bar Chart

```{r ggplot-bar}
ggplot(sales_data, aes(x = Month, y = Sales, fill = Region)) +
  geom_col(show.legend = TRUE, width = 0.7) +
  geom_hline(yintercept = mean(sales_data$Target),
             linetype = "dashed", color = "#e74c3c", linewidth = 0.8) +
  annotate("text", x = 1, y = mean(sales_data$Target) + 5000,
           label = "Avg Target", color = "#e74c3c", hjust = 0, size = 3.5) +
  scale_y_continuous(labels = label_dollar(scale = 1e-3, suffix = "K")) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title    = "Monthly Sales by Region",
    subtitle = "Dashed line = average target",
    x = "Month", y = "Sales", fill = "Region",
    caption  = "Source: Sample dataset"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title    = element_text(face = "bold", color = "#2c3e50", size = 16),
    plot.subtitle = element_text(color = "#7f8c8d"),
    panel.grid.major.x = element_blank(),
    legend.position = "top"
  )
```

## Line Chart with Points

```{r ggplot-line}
ggplot(sales_data, aes(x = seq_along(Month))) +
  geom_ribbon(aes(ymin = Target, ymax = Sales),
              fill = "#1a73e8", alpha = 0.15) +
  geom_line(aes(y = Sales,  color = "Actual Sales"),  linewidth = 1.2) +
  geom_line(aes(y = Target, color = "Target"),  linewidth = 1.2, linetype = "dashed") +
  geom_point(aes(y = Sales,  color = "Actual Sales"),  size = 3) +
  geom_point(aes(y = Target, color = "Target"),  size = 3, shape = 17) +
  scale_x_continuous(breaks = 1:12, labels = month.abb) +
  scale_y_continuous(labels = label_dollar(scale = 1e-3, suffix = "K")) +
  scale_color_manual(values = c("Actual Sales" = "#1a73e8", "Target" = "#e74c3c")) +
  labs(
    title = "Sales vs Target — Monthly Trend",
    x = "Month", y = "Amount (USD)", color = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title  = element_text(face = "bold", color = "#2c3e50"),
    legend.position = "top"
  )
```

## Scatter Plot

```{r ggplot-scatter}
ggplot(sales_data, aes(x = Customers, y = Sales, color = Region, size = Achievement)) +
  geom_point(alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE, color = "#2c3e50",
              fill = "#bdc3c7", linewidth = 0.8, linetype = "dashed") +
  scale_y_continuous(labels = label_dollar(scale = 1e-3, suffix = "K")) +
  scale_size_continuous(range = c(4, 12)) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    title = "Sales vs Number of Customers",
    subtitle = "Point size = Target Achievement %",
    x = "Number of Customers", y = "Sales", color = "Region", size = "Achievement %"
  ) +
  theme_light(base_size = 13) +
  theme(plot.title = element_text(face = "bold", color = "#2c3e50"))
```

---

# plotly — Interactive Charts

`plotly` makes your `ggplot2` charts fully interactive — hover, zoom, pan, click.

## ggplot → plotly Conversion

```{r plotly-ggplotly}
# Build a ggplot first, then wrap with ggplotly()
p <- ggplot(sales_data, aes(
    x    = Month,
    y    = Sales,
    fill = Region,
    text = paste0(
      "Month: ", Month, "<br>",
      "Region: ", Region, "<br>",
      "Sales: $", formatC(Sales, format="f", big.mark=",", digits=0), "<br>",
      "Achievement: ", Achievement, "%"
    )
  )) +
  geom_col(width = 0.7) +
  scale_y_continuous(labels = label_dollar(scale=1e-3, suffix="K")) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Interactive Sales by Month (hover for details)",
       x = "Month", y = "Sales") +
  theme_minimal()

ggplotly(p, tooltip = "text") %>%
  layout(
    hovermode = "x unified",
    legend    = list(orientation = "h", y = -0.2)
  )
```

## Native plotly Chart

```{r plotly-native}
plot_ly(
  data = sales_data,
  x    = ~Month,
  y    = ~Sales,
  name = "Actual Sales",
  type = "scatter",
  mode = "lines+markers",
  line   = list(color = "#1a73e8", width = 2.5),
  marker = list(size = 8, color = "#1a73e8")
) %>%
  add_trace(
    y    = ~Target,
    name = "Target",
    mode = "lines+markers",
    line   = list(color = "#e74c3c", width = 2, dash = "dash"),
    marker = list(size = 7, color = "#e74c3c", symbol = "triangle-up")
  ) %>%
  layout(
    title  = list(text = "Sales vs Target — Interactive Trend", x = 0),
    xaxis  = list(title = "Month"),
    yaxis  = list(title = "Amount ($)", tickformat = "$,.0f"),
    legend = list(orientation = "h", y = -0.2),
    hovermode = "x unified"
  )
```

---

# htmlwidgets — Leaflet Map

`htmlwidgets` is the framework that powers interactive R widgets. `leaflet` is one of the best-known examples.

## Interactive Map

```{r leaflet-map}
# Sample office / regional locations
locations <- data.frame(
  name  = c("HQ - New York", "London Office", "Dubai Office",
             "Singapore Office", "Sydney Office"),
  lat   = c(40.7128, 51.5074, 25.2048, 1.3521, -33.8688),
  lng   = c(-74.0060, -0.1278, 55.2708, 103.8198, 151.2093),
  staff = c(500, 320, 180, 210, 150),
  color = c("red", "blue", "green", "orange", "purple")
)

leaflet(locations) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%       # Clean map style
  addCircleMarkers(
    lat         = ~lat,
    lng         = ~lng,
    radius      = ~sqrt(staff) * 1.5,                   # Size by staff count
    color       = ~color,
    fillOpacity = 0.7,
    popup       = ~paste0(
      "<b>", name, "</b><br>",
      "👥 Staff: ", staff, "<br>",
      "📍 ", round(lat,2), ", ", round(lng,2)
    ),
    label = ~name
  ) %>%
  addLegend(
    position = "bottomright",
    colors   = c("red","blue","green","orange","purple"),
    labels   = locations$name,
    title    = "Office Locations"
  ) %>%
  setView(lng = 20, lat = 25, zoom = 2)
```

<div class="info-box">
🗺️ **Tip:** Click any circle to see the popup. Scroll to zoom. Drag to pan.
Circle size reflects staff count.
</div>

---

# flexdashboard — Dashboard Layout

`flexdashboard` lets you arrange content into side-by-side columns and rows.
Below we simulate the layout using HTML/CSS (since full flexdashboard needs its own output format).

<div class="warning-box">
⚠️ **Note:** Full `flexdashboard` requires `output: flexdashboard::flex_dashboard` in the YAML
header and its own `.Rmd` file. The layout below shows how it looks conceptually.
See the **bonus template** at the bottom of this report for a complete flexdashboard example.
</div>

```{r flexdashboard-gauges}
# Gauge-style metric — requires flexdashboard
# In a real flexdashboard .Rmd these would render as circular gauges
# Here we display the values as an info table

gauge_data <- data.frame(
  Metric  = c("Sales Target", "Customer Satisfaction", "On-Time Delivery", "Budget Used"),
  Value   = c(87, 92, 78, 65),
  Min     = c(0, 0, 0, 0),
  Max     = c(100, 100, 100, 100),
  Symbol  = c("%", "%", "%", "%")
)

kable(gauge_data, caption = "📊 KPI Gauges (values for flexdashboard::gauge())") %>%
  kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE) %>%
  column_spec(2, bold = TRUE,
              background = ifelse(gauge_data$Value >= 80, "#d5f5e3",
                           ifelse(gauge_data$Value >= 60, "#fef9e7", "#fde8e8")))
```

**In a real flexdashboard `.Rmd`, gauges look like this:**

```r
# flexdashboard gauge example (place inside a flexdashboard column)
library(flexdashboard)
gauge(87, min = 0, max = 100, symbol = "%",
      label = "Sales Target",
      gaugeSectors(success = c(80,100), warning = c(60,79), danger = c(0,59)))
```

---

# Custom HTML & CSS Styling

You can embed any HTML directly in R Markdown.

## Callout Boxes

<div class="success-box">
✅ **Success:** All Q4 targets have been met. Sales grew by 18% year-over-year.
</div>

<div class="warning-box">
⚠️ **Warning:** The West region is 12% below target. Review required before month-end.
</div>

<div class="info-box">
ℹ️ **Info:** Next reporting period starts January 1st. Please submit your forecasts by December 20th.
</div>

## Inline R Values in Text

You can embed live R calculations directly in text using backtick-r syntax:

- Total sales this period: **`r paste0("$", formatC(sum(sales_data$Sales), format="f", big.mark=",", digits=0))`**
- Months where target was met: **`r sum(sales_data$Status == "✅ Met")`** out of 12
- Best performing month: **`r sales_data$Month[which.max(sales_data$Sales)]`**
  with **`r paste0("$", formatC(max(sales_data$Sales), format="f", big.mark=",", digits=0))`** in sales

---

# Rendering in Databricks

## Method 1 — Render from R Cell

```r
# Run this in a Databricks R cell to generate the HTML report
rmarkdown::render(
  input       = "/dbfs/your-folder/databricks_dynamic_report.Rmd",
  output_file = "/dbfs/your-folder/output_report.html"
)
```

## Method 2 — Display Inline in Notebook

```r
# Render and show in notebook output
library(IRdisplay)
html_content <- paste(
  readLines("/dbfs/your-folder/output_report.html"),
  collapse = "\n"
)
display_html(html_content)
```

## Method 3 — Save to DBFS and Share

```r
# Copy to a shared DBFS location
file.copy(
  from = "/dbfs/your-folder/output_report.html",
  to   = "/dbfs/FileStore/reports/monthly_report.html"
)

# Access via:
# https://<your-databricks-workspace>/files/reports/monthly_report.html
```

---

# Summary — Package Quick Reference

```{r summary-table, echo=FALSE}
summary_df <- data.frame(
  Package       = c("knitr / kableExtra", "DT", "ggplot2", "plotly",
                    "leaflet", "htmlwidgets", "flexdashboard"),
  Purpose       = c("Styled static tables", "Interactive tables",
                    "Static publication-quality charts",
                    "Interactive hover/zoom charts",
                    "Interactive maps",
                    "Framework for all interactive widgets",
                    "Dashboard layout with gauges & value boxes"),
  `Best For`    = c("Reports, summaries", "Data exploration",
                    "Final presentation charts",
                    "Exploratory & shareable charts",
                    "Geo / location data",
                    "Any widget package", "Executive dashboards"),
  Interactive   = c("No","Yes","No","Yes","Yes","Yes","Yes"),
  check.names   = FALSE
)

kable(summary_df, caption = "📦 Package Summary") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#2c3e50", color = "white") %>%
  column_spec(4, bold = TRUE,
              color = ifelse(summary_df$Interactive == "Yes", "#27ae60", "#e74c3c"))
```

---

*Report generated on `r Sys.time()` | Template version 1.0*
